---
title: "R Notebook for PCA scorings"
output: 
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

# Data extraction, transformation and loading

## Packages loaded

```{r warning=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

#data loading
if (!require("readr")) {
  install.packages("readr", dependencies = FALSE)
  library(readr)

}
citation("readr")

#data manipulation
if (!require("dplyr")) {
  install.packages("dplyr", dependencies = TRUE)
  library(dplyr)
  
}
citation("dplyr")

#data visualisation
if (!require("ggplot2")) {
  install.packages("ggplot2", dependencies = TRUE)
  library(ggplot2)
 
}
citation("ggplot2")

#visuals for pca
if (!require("factoextra")) {
  install.packages("factoextra", dependencies = TRUE)
  library(factoextra)
  
}
citation("factoextra")

#visuals for columntorownames
if (!require("tidyverse")) {
  install.packages("tidyverse", dependencies = TRUE)
  library(tidyverse)
}
citation("tidyverse")

#visuals for columntorownames
if (!require("tidyr")) {
  install.packages("tidyr", dependencies = TRUE)
  library(tidyr)
}
citation("tidyr")
```

## Data loading

```{r}
typeVars <- c("Feet", "Head", "Type","FeatherColor")
ScoringData <- 
  read_delim(
    "../Data/Scorings.csv",
    delim = ";",
    col_types = cols(
      Date = col_character(),
      Sex = col_character(),
      AnimalIdentifier = col_character(),
      Weight = col_integer(), 
      Color = col_character(), 
      Feet = col_integer(), 
      Head = col_integer(),
      Type = col_integer(), 
      FeatherColor = col_integer()
      )) %>%
  dplyr::mutate(
    AverageScore = rowMeans(select(., typeVars))
    ) %>% drop_na() %>%
  dplyr::select(
    AnimalIdentifier,
    Weight,
    Feet, 
    Head,
    Type,
    FeatherColor
  ) %>% column_to_rownames(
    var = "AnimalIdentifier")
```

# PCA

```{r}
pcaScores <- prcomp(ScoringData, scale=TRUE)
```


# Visuals

## Vizualize eigenvalues (scree plot)

Show the percentage of variances explained by each principal component.

```{r}
fviz_eig(pcaScores)
```

## Graph of individual chicken

Individuals with a similar profile are grouped together.

```{r}
fviz_pca_ind(pcaScores,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```

## Graph of variables. 

Positive correlated variables point to the same side of the plot. Negative correlated variables point to opposite sides of the graph.

```{r}
fviz_pca_var(pcaScores,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```

## Biplot of individuals and variables

```{r}
fviz_pca_biplot(pcaScores, repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )
```



